---
title: "LODES Diagnostics --- Neighborhood Level"
format:
    html:
        code-fold: true
---

```{r}
#| echo: false 
#| warning: false
#| output: false
library(tidyverse)
library(sf)

src <- read.csv("../Data/location_characteristics.csv") 
zillow_shp <- st_read("../Data/zillow_neighborhoods",     
        "chicago_zillow_neighborhoods") %>%
    filter(City=="Chicago" & County=="Cook" & 
           Name != "O'Hare International Airport" & 
           Name != "Trumbull Park") %>%
    mutate(regionid = as.numeric(RegionID))

src <- left_join(zillow_shp, src, by="regionid")
```

# Residents and Distribution of Wages Across Origin Neighborhoods

@fig-residents shows the number of residents as summarized by the total number of jobs originating from the particular neighborhood. @fig-homewages shows the distribution of income by where people are commuting from. Note that this is calculated as: $$w_j = \sum_{i=1}^I \pi_{ij} w_i $$
where $\pi_{ij}$ is the probability of commuting from neighborhood $j$ to neighborhood $i$. 
```{r}
#| label: fig-residents
#| fig-cap: Number of Residents
ggplot(src) + 
  geom_sf(aes(fill=num_residents), color="grey70", lwd=0.1) + 
  scale_fill_distiller(palette = "YlOrBr", na.value = "grey50", 
                       direction=1, name = "Population") + 
  theme_void() +
  theme(legend.position = "bottom", legend.box="horizontal",
        legend.key.width = unit(1.5, "cm")) + 
  guides(fill = guide_colorbar(title.position="top", 
                                     title.hjust =0.5))
ggsave("../output/diagnostic_num_residents.pdf")
```

```{r}
#| label: fig-homewages
#| fig-cap: Distribution of Wages
ggplot(src) + 
  geom_sf(aes(fill=income), color="grey70", lwd=0.1) + 
  scale_fill_distiller(palette = "YlOrBr", na.value = "grey50", 
                       direction=1, name = "Income") + 
  theme_void() +
  theme(legend.position = "bottom", legend.box="horizontal",
        legend.key.width = unit(1.5, "cm")) + 
  guides(fill = guide_colorbar(title.position="top", 
                                     title.hjust =0.5))
ggsave("../output/diagnostic_home_wages.pdf")

```

# Employment and Distribution of Wages Across Destination Neighborhoods

@fig-employment shows the number of employees in a certain neighborhood, calculated as the total number of people who are commuting to the neighborhood as a destination. Note that without the log, the loop is very concentrated. This is for aesthetic reasons. @fig-workwages shows the distribution of average wage earned in a particular neighborhood. Clearly, the distribution of the "quality" of jobs are very different compared to the distribution of the "income." While the range here is larger, the concentration of income along the Northeast of Chicago in @fig-homewages is stark.

```{r}
#| label: fig-employment
#| fig-cap: Number of Employees
ggplot(src) + 
  geom_sf(aes(fill=log(num_employment)), color="grey70", lwd=0.1) + 
  scale_fill_distiller(palette = "YlOrBr", na.value = "grey50", 
                       direction=1, name = "Log(Employment)") + 
  theme_void() +
  theme(legend.position = "bottom", legend.box="horizontal",
        legend.key.width = unit(1.5, "cm")) + 
  guides(fill = guide_colorbar(title.position="top", 
                                     title.hjust =0.5))
ggsave("../output/diagnostic_num_employment.pdf")
```


```{r}
#| label: fig-workwages
#| fig-cap: Distribution of Wages At Employment
ggplot(src) + 
  geom_sf(aes(fill=wages), color="grey70", lwd=0.1) + 
  scale_fill_distiller(palette = "YlOrBr", na.value = "grey50", 
                       direction=1, 
                       name = "Wages Offered at Employment Destination") + 
  theme_void() +
  theme(legend.position = "bottom", legend.box="horizontal",
        legend.key.width = unit(1.5, "cm")) + 
  guides(fill = guide_colorbar(title.position="top", 
                                     title.hjust =0.5))
ggsave("../output/diagnostic_work_wages.pdf")
```

# Home Prices

@fig-houseprices shows the distribution of residential house price index. Clearly, the northside has the advantage in residential house prices. Notice that the southwest is particularly cheap. 

```{r}
#| label: fig-houseprices
#| fig-cap: Distribution of House Prices
ggplot(src) + 
  geom_sf(aes(fill=home_price), color="grey70", lwd=0.1) + 
  scale_fill_distiller(palette = "YlOrBr", na.value = "grey50", 
                       direction=1, 
                       name = "Zillow House Price Index") + 
  theme_void() +
  theme(legend.position = "bottom", legend.box="horizontal",
        legend.key.width = unit(1.5, "cm")) + 
  guides(fill = guide_colorbar(title.position="top", 
                                     title.hjust =0.5))
ggsave("../output/diagnostic_prices.pdf")
```

# Commuting gravity
Consider the gravity of people commuting to the loop, which is indexed by the neighborhood ID of 269593. For fun, Hyde Park is 269586. East Hyde Park is 403352. We calculate commuting probability relative to home census tract (the share of commuters commuting to each destination).

```{r}
#| echo: false 
#| warning: false 
#| output: false

commute <- read.csv("../Data/commuting_flow.csv") %>%
    group_by(h_regionid) %>%
    mutate(tot_commute_from_origin = sum(num_commuting)) %>%
    ungroup() %>%
    mutate(share_commuters = num_commuting / 
                tot_commute_from_origin)

zillow_shp <- st_read("../Data/zillow_neighborhoods",     
        "chicago_zillow_neighborhoods") %>%
    filter(City=="Chicago" & County=="Cook" & 
           Name != "O'Hare International Airport" & 
           Name != "Trumbull Park") %>%
    mutate(h_regionid = as.numeric(RegionID))

loop_commute <- commute %>%
    filter(w_regionid == 269593) %>%
    left_join(., zillow_shp, by="h_regionid") %>%
    st_as_sf()

hp_commute <- commute %>%
    filter(w_regionid == 269586) %>%
    left_join(., zillow_shp, by="h_regionid") %>%
    st_as_sf()

ehp_commute <- commute %>%
    filter(w_regionid == 403352) %>%
    left_join(., zillow_shp, by="h_regionid") %>%
    st_as_sf()


cost <- read.csv("../Data/commuting_cost.csv") %>%
    rename(h_regionid = src_id)

loop_cost <- cost %>%
    filter(dst_id == 269593) %>%
    left_join(., zillow_shp, by="h_regionid") %>%
    st_as_sf()

hp_cost <- cost %>%
    filter(dst_id == 269586) %>%
    left_join(., zillow_shp, by="h_regionid") %>%
    st_as_sf()

ehp_cost <- cost %>%
    filter(dst_id == 403352) %>%
    left_join(., zillow_shp, by="h_regionid") %>%
    st_as_sf()

```

Who commutes to good jobs? Those with high income in the income map.

```{r}
ggplot(loop_commute) + 
  geom_sf(aes(fill=share_commuters), color="grey70", lwd=0.1) + 
  scale_fill_distiller(palette = "YlOrBr", na.value = "grey50", 
                       direction=1, 
                       name = "Commuting Probability") + 
  theme_void() +
  theme(legend.position = "bottom", legend.box="horizontal",
        legend.key.width = unit(1.5, "cm")) + 
  guides(fill = guide_colorbar(title.position="top", 
                                     title.hjust =0.5))

ggsave("../output/diagnostic_loop_commutingprob.pdf")
```

How does this relate to commuting cost? Below plots both driving distance and time. At first glance, they seem about equal.

```{r}
ggplot(loop_cost) + 
  geom_sf(aes(fill=duration), color="grey70", lwd=0.1) + 
  scale_fill_distiller(palette = "YlOrBr", na.value = "grey50", 
                       direction=1, 
                       name = "Driving Time") + 
  theme_void() +
  theme(legend.position = "bottom", legend.box="horizontal",
        legend.key.width = unit(1.5, "cm")) + 
  guides(fill = guide_colorbar(title.position="top", 
                                     title.hjust =0.5))
ggsave("../output/diagnostic_loop_commutingcost.pdf")

ggplot(loop_cost) + 
  geom_sf(aes(fill=distance), color="grey70", lwd=0.1) + 
  scale_fill_distiller(palette = "YlOrBr", na.value = "grey50", 
                       direction=1, 
                       name = "Driving Distance") + 
  theme_void() +
  theme(legend.position = "bottom", legend.box="horizontal",
        legend.key.width = unit(1.5, "cm")) + 
  guides(fill = guide_colorbar(title.position="top", 
                                     title.hjust =0.5))
ggsave("../output/diagnostic_loop_commutingcost_distance.pdf")

```

For fun: who commutes to Hyde Park? 
```{r}
ggplot(hp_commute) + 
  geom_sf(aes(fill=share_commuters), color="grey70", lwd=0.1) + 
  scale_fill_distiller(palette = "YlOrBr", na.value = "grey50", 
                       direction=1, 
                       name = "Commuting Probability") + 
  theme_void() +
  theme(legend.position = "bottom", legend.box="horizontal",
        legend.key.width = unit(1.5, "cm")) + 
  guides(fill = guide_colorbar(title.position="top", 
                                     title.hjust =0.5))
ggsave("../output/diagnostic_hp_commutingprob.pdf")

```

How does this relate to commuting cost? Here, when you look at the importance of duration vs. distance, since HP is not central, the cost of commuting looks a bit different. 

```{r}
ggplot(hp_cost) + 
  geom_sf(aes(fill=duration), color="grey70", lwd=0.1) + 
  scale_fill_distiller(palette = "YlOrBr", na.value = "grey50", 
                       direction=1, 
                       name = "Driving Time") + 
  theme_void() +
  theme(legend.position = "bottom", legend.box="horizontal",
        legend.key.width = unit(1.5, "cm")) + 
  guides(fill = guide_colorbar(title.position="top", 
                                     title.hjust =0.5))
ggsave("../output/diagnostic_hp_commutingcost.pdf")

ggplot(hp_cost) + 
  geom_sf(aes(fill=distance), color="grey70", lwd=0.1) + 
  scale_fill_distiller(palette = "YlOrBr", na.value = "grey50", 
                       direction=1, 
                       name = "Driving Distance") + 
  theme_void() +
  theme(legend.position = "bottom", legend.box="horizontal",
        legend.key.width = unit(1.5, "cm")) + 
  guides(fill = guide_colorbar(title.position="top", 
                                     title.hjust =0.5))
ggsave("../output/diagnostic_hp_commutingcost_distance.pdf")

```

What about East Hyde Park?

```{r}
ggplot(ehp_commute) + 
  geom_sf(aes(fill=share_commuters), color="grey70", lwd=0.1) + 
  scale_fill_distiller(palette = "YlOrBr", na.value = "grey50", 
                       direction=1, 
                       name = "Commuting Probability") + 
  theme_void() +
  theme(legend.position = "bottom", legend.box="horizontal",
        legend.key.width = unit(1.5, "cm")) + 
  guides(fill = guide_colorbar(title.position="top", 
                                     title.hjust =0.5))
ggsave("../output/diagnostic_ehp_commutingprob.pdf")

ggplot(ehp_cost) + 
  geom_sf(aes(fill=duration), color="grey70", lwd=0.1) + 
  scale_fill_distiller(palette = "YlOrBr", na.value = "grey50", 
                       direction=1, 
                       name = "Driving Time") + 
  theme_void() +
  theme(legend.position = "bottom", legend.box="horizontal",
        legend.key.width = unit(1.5, "cm")) + 
  guides(fill = guide_colorbar(title.position="top", 
                                     title.hjust =0.5))
ggsave("../output/diagnostic_ehp_commutingcost.pdf")

ggplot(ehp_cost) + 
  geom_sf(aes(fill=distance), color="grey70", lwd=0.1) + 
  scale_fill_distiller(palette = "YlOrBr", na.value = "grey50", 
                       direction=1, 
                       name = "Driving Distance") + 
  theme_void() +
  theme(legend.position = "bottom", legend.box="horizontal",
        legend.key.width = unit(1.5, "cm")) + 
  guides(fill = guide_colorbar(title.position="top", 
                                     title.hjust =0.5))
ggsave("../output/diagnostic_ehp_commutingcost_distance.pdf")

```

```{r}
#| output: false
#| echo: false

blah <- src %>%
      mutate(percentile = round(percent_rank(income) * 100, digits =2)) %>%
      filter(regionid==403352 | regionid==138261)

```

East Hyde Park is in the `r mean(blah[blah$regionid==403352, ]$percentile)`th percentile of the income distribution while Englewood is in the `r mean(blah[blah$regionid==138261, ]$percentile)`th percentile of the income distribution.

# Where do people from Englewood commute to?

```{r}
zillow_shp <- zillow_shp %>%
      rename(w_regionid = h_regionid)

englewood_commute <- commute %>%
    filter(h_regionid == 138261) %>%
    left_join(., zillow_shp, by="w_regionid") %>%
    st_as_sf()

ggplot(englewood_commute) + 
  geom_sf(aes(fill=share_commuters), color="grey70", lwd=0.1) + 
  scale_fill_distiller(palette = "YlOrBr", na.value = "grey50", 
                       direction=1, 
                       name = "Commuting Probability") + 
  theme_void() +
  theme(legend.position = "bottom", legend.box="horizontal",
        legend.key.width = unit(1.5, "cm")) + 
  guides(fill = guide_colorbar(title.position="top", 
                                     title.hjust =0.5))
ggsave("../output/diagnostic_englewood_src_commutingprob.pdf")
```

```{r}
blah <- read.csv("../Data/commuting_flow_bytype.csv") %>%
      group_by(h_regionid) %>%
      summarize(lw = sum(num_commuting_lw),
                hw = sum(num_commuting_hw)) %>%
      mutate(slw = lw / (lw + hw),
             ranking = desc(row_number(slw)) + 226)
```

`r round(mean(blah[blah$h_regionid==138261, ]$slw) * 100, 2)` of the commutes in Englewood are commutes to low-wage jobs (<1500 a month). This ranks `r mean(blah[blah$h_regionid==138261, ]$ranking)`th across all neighborhoods.

# Where do people from East Hyde Park commute to?
```{r}
ehp_commute <- commute %>%
    filter(h_regionid == 403352) %>%
    left_join(., zillow_shp, by="w_regionid") %>%
    st_as_sf()

ggplot(ehp_commute) + 
  geom_sf(aes(fill=share_commuters), color="grey70", lwd=0.1) + 
  scale_fill_distiller(palette = "YlOrBr", na.value = "grey50", 
                       direction=1, 
                       name = "Commuting Probability") + 
  theme_void() +
  theme(legend.position = "bottom", legend.box="horizontal",
        legend.key.width = unit(1.5, "cm")) + 
  guides(fill = guide_colorbar(title.position="top", 
                                     title.hjust =0.5))
ggsave("../output/diagnostic_ehp_src_commutingprob.pdf")
```

`r round(mean(blah[blah$h_regionid==403352, ]$slw) * 100, digits=2)` of the commutes in East Hyde Park are commutes to low-wage jobs (<1500 a month). This ranks `r mean(blah[blah$h_regionid==403352, ]$ranking)`th across all neighborhoods.

# Where do people from Lincoln Park commute to?

```{r}
lp_commute <- commute %>%
    filter(h_regionid == 269590) %>%
    left_join(., zillow_shp, by="w_regionid") %>%
    st_as_sf()

ggplot(lp_commute) + 
  geom_sf(aes(fill=share_commuters), color="grey70", lwd=0.1) + 
  scale_fill_distiller(palette = "YlOrBr", na.value = "grey50", 
                       direction=1, 
                       name = "Commuting Probability") + 
  theme_void() +
  theme(legend.position = "bottom", legend.box="horizontal",
        legend.key.width = unit(1.5, "cm")) + 
  guides(fill = guide_colorbar(title.position="top", 
                                     title.hjust =0.5))
ggsave("../output/diagnostic_lp_src_commutingprob.pdf")
```

`r round(mean(blah[blah$h_regionid==269590, ]$slw) * 100, digits = 2)` of the commutes in Lincoln Park are commutes to low-wage jobs (<1500 a month). This ranks `r mean(blah[blah$h_regionid==269590, ]$ranking)`th across all neighborhoods.

# These are remarkably similar: all of the gravity in Chicago in terms of commuting probability is in downtown. What kind of jobs come from where? 

```{r}
#| echo: false 
#| output: false

zillow_shp <- st_read("../Data/zillow_neighborhoods",     
        "chicago_zillow_neighborhoods") %>%
    filter(City=="Chicago" & County=="Cook" & 
           Name != "O'Hare International Airport" & 
           Name != "Trumbull Park") %>%
    mutate(h_regionid = as.numeric(RegionID))

```
```{r}
#zillow_shp <- zillow_shp %>%
#      rename(h_regionid = w_regionid)

commute_by_type <- read.csv("../Data/commuting_flow_bytype.csv") %>%
    group_by(w_regionid) %>%
    mutate(total_commute = sum(num_commuting_lw + num_commuting_hw),
           total_commute_lw = sum(num_commuting_lw)) %>%
    ungroup() %>%
    mutate(share_low_wage_commuters = num_commuting_lw / 
                total_commute_lw)
loop_lw <- commute_by_type %>%
    filter(w_regionid == 269593) %>%
    left_join(., zillow_shp, by="h_regionid") %>%
    st_as_sf()

```

@fig-shareloop_lowwage_commute shows the commuting origins of low-wage workers in the loop.

```{r}
#| label: fig-shareloop_lowwage_commute
#| fig-cap: Low-wage Commuters to the Loop

ggplot(loop_lw) + 
  geom_sf(aes(fill=share_low_wage_commuters), color="grey70", lwd=0.1) + 
  scale_fill_distiller(palette = "YlOrBr", na.value = "grey50", 
                       direction=1, 
                       name = "Commuting Probability") + 
  theme_void() +
  theme(legend.position = "bottom", legend.box="horizontal",
        legend.key.width = unit(1.5, "cm")) + 
  guides(fill = guide_colorbar(title.position="top", 
                                     title.hjust =0.5))
ggsave("../output/diagnostic_loop_lw_commutingprob.pdf")
```

# More generally, where do these low wage workers come from in the city? 

@fig-sharelowtypebyhome shows that the low-wage workers in Chicago come disproportinately from the south and southwest of Chicago. 


```{r}
blah <- read.csv("../Data/commuting_flow_bytype.csv") %>%
      group_by(h_regionid) %>%
      summarize(lw = sum(num_commuting_lw),
                hw = sum(num_commuting_hw)) %>%
      mutate(slw = lw / (lw + hw),
             ranking = desc(row_number(slw)) + 226) %>%
      left_join(., zillow_shp, by='h_regionid') %>%
      st_as_sf()
```

```{r}
#| label: fig-sharelowtypebyhome
#| fig-cap: Share of Commuters that Earn Low Wages
ggplot(blah) + 
  geom_sf(aes(fill=slw), color="grey70", lwd=0.1) + 
  scale_fill_distiller(palette = "YlOrBr", na.value = "grey50", 
                       direction=1, 
                       name = "Share of Commuters that are Low-Wage") + 
  theme_void() +
  theme(legend.position = "bottom", legend.box="horizontal",
        legend.key.width = unit(1.5, "cm")) + 
  guides(fill = guide_colorbar(title.position="top", 
                                     title.hjust =0.5))
ggsave("../output/diagnostic_city_share_lw.pdf")

```

For exploration:
```{r}
mapview::mapview(zillow_shp)
```